---
title: "Lab8"
author: "put your name here"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Real data example

Wine Quality Data Set is quite interesting dataset with 12 variables. The goal is to predict the quality of wine based on some physicochemical properties.

Read more about this dataset from here: https://archive.ics.uci.edu/ml/datasets/Wine+Quality

Load the dataset to R.

```{r}
url <- 'https://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/winequality-white.csv'
wine <- read.table(url, sep=";", header=TRUE)
head(wine)
```

## Fitting

Now, let's fit following models (most of them can be fitted with caret):

* standard linear regression with all variables (with `caret` or function `lm`),
* choose best models according to AIC/BIC criteria (useful functions: `bincombinations{e1071}` for all subsets, `AIC`/`BIC`/`logLik` for calculation of BIC/AIC, `step` for stepwise regression),
* choose best models according to ridge regression/lasso methods (try `glmnet{glmnet}` and `cv.glmnet` for cross validation. Find more about these functions here: https://web.stanford.edu/~hastie/glmnet/glmnet_alpha.html),
* choose best model with PCR regression (try `pcr{pls}`).

Note that for some of these modes you need to do some parameter tuning.
Do this!
Write a function that will automate this parameter tuning for you.

## Performance on the training set

For each model calculate the MSE (mean square error) on the test set.

## Performance on the test set

It's time to assess the performance of these models. 
Let's divide the dataset into two groups: training (3/4 of data) and testing (1/4) dataset. You can try different proportions, the data is large so it should not matter. The splitting can be done with  `createDataPartition{caret}`.

1. Fit all models listed above on the training dataset.
2. Test their performance on test dataset. Calculate MSE (mean square error) on test set.
3. Repeat the splitting 10 - 100 times (depending on how fast is your solution. Maybe it's feasible to do this 100 times). 
4. Plot the distribution of MSE for different methods (vioplots/boxplots/histograms). 

Which method gives the lowest average MSE in this case?

